---
title: "Class 10 Notes and Exercises"
author: ZN
date: "Date created: `r Sys.Date()`"
output: html_notebook
---

*LEARNING OBJECTIVE*: Explore data effectively using a checklist
*TOPICS TO COVER*

1. Next assignment: Explore the causes of departure and arrival delays
2. Go through the exploratory analysis checklist
   + Formulate your initial question
   + Examine your data sources
       + `str()`, `glimpse()`, `summary()`, `skim()`
       + Check the top and bottom of your data
3. Examine the variables of interest for your question
4. Explore a solution for your question
   + Use visual comparisons
   + Illustrate causal frameworks
   + Use multivariate comparisons
5. Try and invalidate your solution
6. Reformulate your question (Rinse and repeat)
7. Integrate different modes of presenting evidence

--
```{r}
library(tidyverse); library(nycflights13)
```


##Task 1: Examine the weather data
1. Use the `skim()` function from the `skimr` package to visually display all the variables in weather
```{r}
#install.packages("skimr")
library(skimr)
weather 
  skim(weather)
```


2. Use `nrow()` to display the number of rows in the weather data
```{r}
weather %>% 
  nrow()
```

3. Use `dim()` to display the rows and columns of the weather data

```{r}
weather %>% 
  dim()
```

4. Use `names()` to display the column names of the weather data
```{r}
weather %>% 
  names()
```


5. Display the `head()` of the weather data
```{r}
weather %>% 
  head()
```


6. Display the `tail()` of the weather data
```{r}
weather %>% 
  tail()
```


7. Display a random 10 rows from the weather data using `sample_n()`
```{r}
weather %>% 
  sample_n(size = 10)
```

8. What are the primary keys of the weather data? i.e. which column(s) can be used to uniquely identify each observation in the weather data? Hint: `group_by()` with `filter(n() > 1)` to check if there are more than one observation per group.
```{r}
weather %>% 
  group_by(time_hour, origin) %>% 
  #summarise(nobs = n()) #This is just to explore how many obs (count) per group
  filter(n() > 1)
```

```{r}
weather
```

##Task 2: Explore the indepedent variable
Use a combination of text, tables and charts to explore the nature of the independent variable
```{r}
weather %>% 
  summary(wind_gust)
```
```{r}
##histogram
weather %>% 
  ggplot(., aes(wind_gust)) +
    geom_histogram(bins = 100)
```

```{r}
##density plot
weather %>% 
  ggplot(aes(wind_gust)) +
    geom_density()
#This is the showing that around 8% of obs take on the value 20 to 30 mph wind gust
```


```{r}
weather %>% 
    filter(wind_gust > 0) %>% 
    mutate(binnedWindgust = cut_width(wind_gust, width = 10)) %>% 
    group_by(binnedWindgust) %>% 
    summarise(nObs = n())
```

##Task 3. Explore a relationship
Explore the relationship between departure delay and a variable of your choice.


```{r}
weatheredFlights <- flights %>% 
  select(origin, time_hour, dep_delay) %>% 
  left_join(weather, by = c("origin", "time_hour")) 
  weatheredFlights
```



```{r}
#Let's do a few plots
weatheredFlights %>% 
  filter(dep_delay > 0) %>% 
  group_by(wind_gust) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
  ggplot(aes(wind_gust, meanDelay))+
  geom_point()
```


