---
title: "Class 16 Notes and Exercises"
author: ZN
date: "Date created: `r Sys.Date()`"
output: html_notebook
---

###Task Set 1: Conditionals

The code below samples 10 values at random from a vector containing the values from 1 to 30 and 5 missing values (generated by the `rep()` command). Run the entire chunk below to generate the vector x.

```{r}
x <- sample(c(1:30, rep(NA, times = 5)), size = 10)
x
```

1. Check if there are `any()` missing values in x using `is.na()`
```{r}
any(is.na(x))
```


2. If there are `any()` missing values in x then replace all values in x with its squared value else replace it with the square root. You can calculate the square using the following syntax `x ^ 2` and the square root using `x ^ 0.5`
```{r}
if(any(is.na(x))) {x^2} else {x^0.5}
```


3. Use `all()` to check if all the values in x are numeric. You can check if a value is numeric using `is.numeric()`
```{r}
all(is.numeric(x))
```


4. If `all()` values in x are numeric then replace all values in x with its squared value else replace it with the square root.
```{r}
if(all(is.numeric(x))) {x^2} else {x^0.5}
```

5. Use an `if_else()` here. If x is NA then replace it with 1 else replace it with the square of x. Make sure you have the tidyverse loaded since `if_else()` comes from dplyr. The syntax for an `if_else()` is as follows: `if_else(condition, val1, val2)` where condition is something evaluates to a TRUE or FALSE vector that is same length as the val1 and val2.
```{r}
if_else(is.na(x), 1, x^2)
```


6. Use `case_when()` here. If x is NA then replace it with the string "missing". If it is divisible by 2 (i.e. x %% 2 == 0) then replace it with "even". Everything else should be labelled "odd". The syntax for `case_when()` is as follows: `case_when(condition1 ~ val1, condition2 ~ val1,...., TRUE ~ valueforeverythingelse)`

