---
title: "Class 1 Exercises"
output: html_notebook
---

# My first exercise

1. Create the folder structure for the class on your computer called "yourfirstnameR4D" with the following subfolders in it - "Assignments", "Classes" and "Data". This will be your folder for the entire course.
2. Create an R notebook change its YAML title to "Class 1 Exercises"
3. Create a level 2 header called "My first exercise"
4. Create your first code chunk (remember to use the shortcuts)
5. Within it assign any value you would like to a variable. Print this value. Save the file in the "Classes" subfolder
6. Preview the notebook.
7. Use this notebook to take in class notes, try out things and complete in class tasks.
```{r}
classFirstVar <- "Rad"
classFirstVar
```
# My Second Exercise
1. Create and store a character vector (of your choice), a numeric vector (interger/double), all of the same length
2. Use `seq()` to generate an atomic vector containing all the odd numbers less than 100
3. Store all the vectors created in step 1 in a named list
4. Create a dataframe using all the vectors stored in step 1
5. Predict the output of the following without running the code in `c("FALSE", FALSE)`, `c(1.0, FALSE)`, `c(2, TRUE)` and `c("two", 2)`
6. What get returned when you check the `typeof()` the dataframe created in step 4. Why?

```{r}
IceCreamFlavors <- c("Vanilla", "Chocolate", "Mango")
IceCreamFlavors
Price <- c(5, 7, 6)
Price
oddnumseq <- seq(from = 1, to = 99, by = 2)
oddnumseq
vars <- list(IceCreamFlavors, Price)
vars
Icecream <- data.frame(Flavor = IceCreamFlavors, Price = Price)
data.frame
```

