---
title: "Class 4 Notes and Exercises"
author: ZN
date: "Date created: `r Sys.Date()`"
output: html_notebook
---
*Learning Objectives*
1. Recap and learn the importance of ungrouping
2. Create and understand factor variables
3. Understand the need for visualizing data (without summarizing it) before analyzing it.
4. Create simple charts using the layered grammer of graphics
   + A barplot
   + A histogram
   + A column chart
   + A line chart
   + A scatter plot
5. Make facetted plots
6. Draw multiple geoms in one plot
7. Learn to make beautiful charts using
   + Aesthetics (alpha, colour etc.)
   + Scales
   + Theme
   + Labels and annotations
   + Extensions: ggthemes and others
8. Save a plot
8. Build a complex chart together

--
## Recap of last week
The different dplier verbs we've covered so far: The top 5 are most important
1. arrange
2. filter
3. mutate
4. summarize
5. select
6. slice
7. group_by
8. pull

Lets select, filter, mutate a range in a single piped codechunk
```{r, message=FALSE}
library(tidyverse); library(nycflights13)

flights %>% 
  select (1:3, contains("delay"), carrier, 7) %>% 
  filter(arr_time < 1200) %>% 
  mutate(newDate = paste(year, month, day, sep = "-")) %>% 
  arrange (desc(month))
```
 
 Let's look at grouping and summarize
```{r}
flights %>%
  group_by(month) %>% 
  summarize(meanArrDelay = mean(arr_delay, na.rm = T))
```
 
Anytime you group a tibble it is important to ungroup it
```{r}
aGrpflight <- flights %>% 
  group_by(origin)
glimpse(aGrpflight)
```


#LESSON 4 Exercises:
**CHEATSHEETS**

1. [Data visualization](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)
2. [Data Manipulation](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)

**KEYBOARD SHORTCUTS**:

1. Assignment operator (<-) is `Alt+-` for Windows and `Option+-` for Mac
2. Insert new code chunk `Ctrl+Alt+I` for Windows and `Command+Option+I` for Mac
3. Run a line of code `Ctrl+Enter` for Windows and `Command+Enter` for Mac
4. Run all the code within a chunk `Ctrl+Shift+Enter` for Windows and `Command+Shift+Enter` for Mac
5. Insert a pipe operator (%>%) `Ctrl+Shift+M` for Windows and `Command+Shift+M` for Mac

##Task Set 1: Recap + Factors
1. Filter all departure delays that were between 10 and 20 minutes from LGA in 2013
```{r}
library(tidyverse); library(nycflights13)
flights %>%
  select(1, 2, 3, 4, 5, contains("dep"), origin) %>% 
  filter(dep_delay >=10 & dep_delay <=20 & origin == "LGA" & year == "2013")
  
```


2. Use `group_by()` and `filter()` to find the flights that had the greatest delay for each month in 2013.
```{r}
flights %>%
  group_by(month) %>% 
    filter(dep_delay == max(dep_delay, na.rm = T)) %>% 
    arrange(month)
  # this is a logical operation
```

3. Write two lines on why it is important to ungroup a grouped dataset (once your are done with the grouping operations).
 - Ans: so as to avoid forgetting about the group at a later time.
 
 
4. Look at the documentation for `weekdays()`. Use it on the `time_hour` column inside a `mutate()` to create a new column called `dayName` that records the day of the week that a flight took off. Save this new dataset as `flightsWeek`.
```{r}
flightsweek <- flights %>% 
  mutate(dayName = weekdays(time_hour)) #%>% 
  #group_by(dayName) %>% 
  #summarize(meanDelay = mean(dep_delay, na.rm = T))
```

## Let's factor
factor variables are the way in which R handles categorical vars
```{r}
genderVar <- sample(x = c("M","F", "Other"), size = 200, replace = T)
genderVar
table(genderVar)
```

```{r}
fctGender <- factor(genderVar)
str(fctGender)
levels(fctGender) <- c("Other", "F", "M") # this is setting levels on how you're ordering the table
table(fctGender)
```
or you can do it like this
```{r}
fctGenderFinal <- factor(genderVar, levels = c("Other", "F", "M"), labels = c("Other", "F", "M"))
table(fctGenderFinal)
```

5. Convert this into a factor variable called `fctDayNames` using `factor(flightsWeek$dayName)`. Display it using `table(fctDayNames)`
```{r}
fctDayNames <- factor(flightsweek$dayName)
table(fctDayNames)
```

6. Use `levels()` on `fctDayNames` to order the levels of the factors to the following order: c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"). Show the output using `table(fctDayNames)`
```{r}
levels(fctDayNames) <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
table(fctDayNames)
                  
```

7. Now combine the previous two steps in a `factor()` function by setting the value of the levels argument within it, also set the labels argument to c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"). Save this as `cleanFctDayNames` See `?factor`. Show output using `table(cleanFctDayNames)`. Observe the differences in output between 5, 6 and 7.

```{r}
cleanFctDayNames <- factor(fctDayNames, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"), labels = c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
table(cleanFctDayNames)
```

# Back to learning objectives - GG plots

Lets start with a basic histogram
```{r}
flights %>% 
  filter(dep_delay < 250) %>% 
  ggplot(data = ., mapping = aes(dep_delay)) +
  geom_histogram(bins=100)
```

Let's look at a barplot
```{r}
ggplot(data = flights, mapping = aes(carrier)) +
  geom_bar()
```
Let's make a column chart (barchart)
```{r}
flights %>%
  group_by(carrier) %>% 
  summarize(meanDelay = mean(dep_delay, na.rm = T))%>% 
  ggplot(aes(x = carrier, y = meanDelay)) +
  geom_col()

```

##Task Set 2: Create simple charts
1. Create a histogram of arrival delay.
```{r}
flights %>% 
  ggplot(data = ., mapping = aes(arr_delay)) +
  geom_histogram(binwidth=5)
```

2. Change the statistic for the histogram from the default of "count" to "density" and plot again.
```{r}
flights %>% 
  ggplot(data = ., mapping = aes(arr_delay)) +
  geom_histogram(stat = "density")
```

3. Run the following code to create a stacked column chart. Look through the documentation and examples for `geom_col()` on the tidyverse portal to fix the code below and create a chart where the bars for each origin are not stacked.
```{r}
flights %>%
    group_by(origin, month) %>%
    summarise(meanDelay = mean(dep_delay, na.rm = T)) %>%
    ggplot(aes(x = factor(month), y = meanDelay, fill = origin)) +
    geom_col()
```


4. Read the documentation for `geom_point()` to create a scatter plot that compares arrival delay with departure delay. How would you describe the relationship between the two variables?
```{r}

```

5. Use the chunk below to group the data by `time_hour` and calculate the mean departure delay and pipe to a ggplot to create a line chart showing `time_hour` on the x-axis and the mean departure delay on the y-axis. Can you identify the spikes in the average delay? Which parts of this would you zoom into to get a better sense for the average delay? 
```{r, fig.width=12}

```


