---
title: "Assignment 2"
author: "Zeenat Nazir"
date: "Date created: `r Sys.Date()`"
output: html_notebook
---
**CHEATSHEETS**

1. [Data visualization](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)
2. [Data Manipulation](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)


Create a new notebook (copy parts of the YAML header you are interested in keeping) to work on your assignment.
The assignment should be submitted via a shareable github link that points to your notebook (.nb.html) file to me on Slack.
Make sure the notebook is neatly formatted, and that all the charts and tables and look good when you preview it.
Task: Explore at least 5 different causes for departure and arrival delays (combined - not 5 each). Its not necessary that you find a definitive answer for what causes a delay, instead I want you to showcase your ability to explore the data with a combination of analytical text, tables and charts. The final output should be something you can show a prospective employer or client as proof of you ability to work with data in R.

## Loading and installing relevant packages for this assignment
```{r}
library(tidyverse); library(nycflights13); library(skimr); library(timeDate); library(ggthemes) 
#install.packages("timeDate")
#install.packages("lubridate")
library(lubridate)
```
##Viewing all the datasets in nycflights

```{r}
flights
```

```{r}
weather 
  #skim(weather)
```
```{r}
airports
#skim(airports)
```

```{r}
airlines
```

```{r}
planes
```

## Merging the weather, airlines and flights datasets. 
####*Ignoring airports for the purpose of this assignment. Starting with a full merged dataset (and narrow down later by hypothesis).

### We start first by exploring keys for different datasets
```{r}
#This is exploring keys to merge weather to flights and the other relevant datasets
weather %>% 
  group_by(time_hour, origin) %>% 
  #summarise(nobs = n()) #This is just to explore how many obs (count) per group
  filter(n() > 1)
```
```{r}
#This is exploring keys to merge planes to flights and the other relevant datasets
planes %>% 
  group_by(tailnum) %>% 
  #summarise(nobs = n()) #This is just to explore how many obs (count) per group
  filter(n() > 1)
```

### Merging weather and airlines and planes to flights with established keys
```{r}
weatheredFlights <- flights %>% 
  left_join(weather, by = c("origin", "time_hour")) %>% 
  left_join(airlines, by = c("carrier")) %>% 
  left_join(planes, by = c("tailnum"))
  weatheredFlights
```

## Formulating initial hypotheses for causes of arrival and departure delays
+ **Question 1: Is there a relationship between the carrier you fly and delays. That is, do certain airlines have a track record of departure delays?**

### Examining data sources and independent var
```{r}
weatheredFlights %>% 
names()
```

### Examining how many observations (flights) per airline
```{r}
weatheredFlights %>% 
  filter(dep_delay > 0) %>% 
  group_by(name) %>% 
  count(name) %>% 
  arrange(n)
```

### Examining avg departure delay by carrier
```{r}
weatheredFlights %>% 
  filter(dep_delay > 0) %>% 
  group_by(name) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
  arrange(desc(meanDelay)) %>% 
  ggplot(aes(x = reorder(name, meanDelay), y = meanDelay)) +
  geom_col(stat = 'identity') +
  coord_flip() +
    labs(
    x = "Airline",
    y = "Average Departure Delay (in minutes)",
    title = "Regional airlines tend to have higher departure delays",
    subtitle = "Among the larger domestic airlines, United's avg delayed departures are lowest"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```




+ **Question 2: Which days (if any) of the week have higher departure delays and if so, which airports have relatively high dep delays?**

### Identifying and using the right variable to extract day of the week. And then summarising dep delays by day of the week. Question for Hari - how do I reverse the appearance of the legend such that LGA is top and EWR bottom to more cleanly align with the bars?
```{r}
weatheredFlights %>% 
mutate(date = as.Date(time_hour)) %>% 
mutate(dayofweek = wday(date, label = TRUE)) %>% 
  filter(dep_delay > 0) %>% 
  group_by(dayofweek, origin) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
  arrange(desc(meanDelay)) %>% 
  ggplot(aes(dayofweek, meanDelay, fill = origin)) +
   geom_bar(stat = 'identity', position=position_dodge()) +
   geom_text(aes(label= round(meanDelay)), vjust=0.4, color="black", hjust = -0.05, 
            position = position_dodge(1), size=3.2) +
  scale_fill_brewer(palette= "Paired") +
    coord_flip() +
    labs(
    x = "Day",
    y = "Average Departure Delay (in minutes)",
    title = "Skip the queues, fly on weekends!"
    ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```


+ **Question 3: Does visibility (along with precipitation) affect dep delays?**

### Exploring the visibility var
```{r}
weather %>% 
  ggplot(aes(visib)) +
  geom_histogram(binwidth = 1)
```
### Exploring the precip var
```{r}
weather %>% 
  ggplot(aes(precip)) +
  geom_histogram(binwidth = 0.05)
```


### Exploring how visibility alone relates to delay
```{r}
#Explore how visibility relates to delay
weatheredFlights %>% 
  filter(dep_delay > 0 & !is.na(dep_delay) & !is.na(visib)) %>% 
  group_by(visib) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
  ggplot(aes(visib, meanDelay))+
  geom_point() +
    geom_smooth(method = "lm") +
  #panel.background = element_rect(fill = NA) +
    labs(
    x = "Visibility (Feet)",
    y = "Average Departure Delay (minutes)",
    title = "Lower visibility translates to higher departure delays"
    ) +
  theme_economist() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

```

### Exploring how visibility along with precipitation relates to delay. Question for Hari - how do I add label to visib? Also, how do I get both visib groups into the same chart with different colors and lines.
```{r}
#Let's explore how precip and visibility relate to delay
weatheredFlights %>% 
  filter(dep_delay > 0 & !is.na(dep_delay) & !is.na(temp) & !is.na(visib)) %>% 
  mutate(visibGroup = cut_interval(visib, length = 5, na.rm = T)) %>%
  group_by(precip, visibGroup) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T)) %>% 
  ggplot(aes(precip, meanDelay))+
  geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(~visibGroup) +
      labs(
    x = "Precipitation (in inches)",
    y = "Average Departure Delay (in minutes)",
    title = "As rain and fog increases, so do delays",
    subtitle = "Visibility"
    ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

```


+**Question 4: Does the volume of flights on any given day affect delays?**
```{r}
depdatemeandel <- weatheredFlights %>% 
  mutate(depDate = paste(year.x, month.x, day.x, sep = "-")) %>% 
  group_by(depDate) %>% 
  mutate(flightVol = n()) %>% 
  filter(dep_delay > 0 & !is.na(dep_delay) & !is.na(depDate)) %>% 
  summarise(meanDelay = mean(dep_delay, na.rm = T))
depdatemeandel

  
```
```{r}
flightvol <- weatheredFlights %>% 
  mutate(depDate = paste(year.x, month.x, day.x, sep = "-")) %>% 
  group_by(depDate) %>% 
  mutate(flightVol = n()) 
flightvol
```

```{r}
newqfourtibble <- depdatemeandel %>% 
  _join(flightvol, by = c("depDate"))
newqfourtibble
```


```{r}
ggplot(aes(flightVol, meanDelay))+
  geom_point() +
    geom_smooth(method = "lm") +
    #facet_wrap(~visibGroup) +
      labs(
    x = "Number of Flights Per Day",
    y = "Average Departure Delay (in minutes)",
    title = "??"
    ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
```


*Question 5: 


```{r}
weatheredFlights
```


------
1. Next assignment: Explore the causes of departure and arrival delays
2. Go through the exploratory analysis checklist
   + Formulate your initial question
   + Examine your data sources
       + `str()`, `glimpse()`, `summary()`, `skim()`
       + Check the top and bottom of your data
3. Examine the variables of interest for your question
4. Explore a solution for your question
   + Use visual comparisons
   + Illustrate causal frameworks
   + Use multivariate comparisons
5. Try and invalidate your solution
6. Reformulate your question (Rinse and repeat)
7. Integrate different modes of presenting evidence
------


